---
title: "Getting Started"
css: ../css/surveydown.css
---

```{r}
#| echo: false
#| eval: true

library(surveydown)
shiny::shinyOptions(bootstrapTheme = bslib::bs_theme(version = 5L))
```

## Overview

{{< include ../chunks/files.qmd >}}

The [{surveydown}]({{< var url_package >}}) R package provides a set of functions for defining the survey content and configuration options. Each function starts with `sd_` to make them easy to identify.

The platform is based on some basic principles:

- Add content to your `survey.qmd` file using markdown text (or in RStudio use the visual editor).
- Define survey questions in R code chunks with the `sd_question()` function.
- Define pages using fences (`:::`), with navigation buttons handled using the `sd_next()` function.
- Add rich functionality to your survey using a variety of [server options](server-options.qmd) and [conditional control logic](conditional-control.qmd) in the `server()` function in the `app.R` file.
- Store your respondent data in a database (see [Storing Data](storing-data.qmd)).

This approach ensures a flexible survey platform that is fully reproducible and easy to customize.

The remaining steps on this page will guide you through the process of creating a surveydown survey.

## 1. Install

See the [Installation](installation.qmd) page.

## 2. Start with a template

::: {.callout-important}

Every survey created with {surveydown} should be in its own separate project folder.

:::

We recommend starting with a template to build your surveydown survey.

In the R console, run the following to to setup a template survey:

```{r}
surveydown::sd_create_survey("path/to/folder")
```

This will create a folder located at `"path/to/folder"` with the following files:

- `survey.qmd`: a template survey where you define your survey content (text, questions, images, etc.).
- `app.R`: a template app file where you define your database connection, survey configuration options, and any other server logic.
- `survey.Rproj`: An RStudio project file (helpful if you're working in RStudio).

## 3. Add survey content in your `survey.qmd` file

See the [Survey Components](survey-components.qmd) page for details on the main components in a surveydown survey. For a quick overview, here's how you add pages and questions:

- Add pages with fences, like this:

```{r}
::: {#page1 .sd-page}

Page 1 content here

:::
```

- Add questions with the `sd_question()` function in code chunks (see the [Question Types](question-types.qmd) page for more on the types of questions supported). For example, here's a multiple choice question:

::: {.panel-tabset}

## Code chunk

```{r}
#| echo: fenced

sd_question(
  type  = 'mc',
  id    = 'penguins',
  label = "Which is your favorite type of penguin?",
  option = c(
    'Adélie'    = 'adelie',
    'Chinstrap' = 'chinstrap',
    'Gentoo'    = 'gentoo'
  )
)
```

## Output

```{r}
#| eval: true
#| echo: false

sd_question(
  type  = 'mc',
  id    = 'penguins',
  label = "Which is your favorite type of penguin?",
  option = c(
    'Adélie'    = 'adelie',
    'Chinstrap' = 'chinstrap',
    'Gentoo'    = 'gentoo'
  )
)
```

:::

## 4. Add control options

In the `server()` function in the `app.R` file, add rich functionality to your survey using a variety of [server options](/server-options.qmd) and [conditional control logic](conditional-control.qmd).

## 5. Setup your database

In the global settings at the top of the `app.R` file, setup your database with the `sd_database()` function. You can also leave it blank to preview / edit your survey without database connected, or set `ignore = TRUE` to run the survey without storing data. See the [Storing Data](storing-data.qmd) page for more details.

## 6. Locally preview

Preview your survey by clicking the "Run App" button in RStudio or in your R console running the `runApp()` command.

## 7. Deploy

Deploy your survey by hosting it on your favorite server, like {{< var shinyapps >}}, {{< var huggingface >}}, [Posit Connect Cloud](https://connect.posit.cloud/), [Heroku](https://www.heroku.com/), etc. See the [Deployment](deployment.qmd) page for more details.
