---
title: "External Redirect"
css: ../css/surveydown.css
---

```{r}
#| eval: true
#| include: false

library(surveydown)
shiny::shinyOptions(bootstrapTheme = bslib::bs_theme(version = 5L))
```

Like other operations, there are also two types of redirection: static and reactive. A **static** redirect provides a hard-coded URL, like [google.com](https://www.google.com), [apple.com](https://www.apple.com), etc. A **reactive** redirect grabs parameters from your survey URL and allows you to use them to construct a new redirect URL.

We recommend you try out our [demo survey](https://github.com/surveydown-dev/demos/tree/main/external-redirect) that showcases both static and reactive redirects so that you can have a full understanding of how this works:

## Static Redirect

In your **survey body**, use the `sd_redirect()` function and define your parameters. Below is an example of a redirect button that goes to [google.com](https://www.google.com).

::: {.panel-tabset}

## Code chunk

```{r}
# Static redirect
sd_redirect(
  id     = "redirect",
  url    = "https://www.google.com",
  label  = "Redirect to Google",
  button = TRUE,
  newtab = TRUE
)
```

## Output

```{r}
#| eval: true
#| echo: false

sd_redirect(
  id     = "redirect",
  url    = "https://www.google.com",
  label  = "Redirect to Google",
  button = TRUE,
  newtab = TRUE
)
```

:::

## Reactive Redirect

Reactive redirect is essentially useful if you deploy your survey on survey panels like [Prolific](https://www.prolific.com) or [Dynata](https://www.dynata.com). These survey panels usually use different IDs to identify survey respondents.

To make reactive redirect work, you need to edit both in your **server chunk**, and your **survey body**. We'll use an example to explain the process. This is the same example in our [demo survey](https://github.com/surveydown-dev/demos/tree/main/external-redirect).

**Firstly**, in your **server chunk**, define `url_normal`. It is an [reactive expression](https://shiny.posit.co/r/getstarted/shiny-basics/lesson6/) in Shiny apps. You'll use the `sd_get_url_pars()` function to retrieve all the parameters in your survey URL, and concatenate with customized texts including id names, status values, etc. In this example, we defined 3 IDs: `id_a`, `id_b`, and `id_c`, which should be grabbed from your survey link.

```{r}
# Reactive expression that generates a url for a normal ending
url_normal <- reactive({
  params <- sd_get_url_pars()
  id_a <- params["id_a"]
  id_b <- params["id_b"]
  id_c <- params["id_c"]
  return(paste0(
    "https://www.google.com?id_a=", id_a,
    "&id_b=", id_b,
    "&id_c=", id_c,
    "&status=0" # status of 0 indicates normal ending
})
```

**Secondly**, also in your **server chunk**, create the redirect button using `sd_redirect()`. Note that we passed `url_normal()` value, with a pair of **parentheses**, to the `url` parameter.

```{r}
# Create the redirect button for normal ending
sd_redirect(
  id = "redirect_normal",
  url = url_normal(),
  button = TRUE,
  label = "Redirect with Normal Status"
)
```

::: {.callout-tip}
You may have noticed that we created `url_normal`, and then passed `url_normal()` to `sd_redirect()`. This is how Shiny app deals with reactivity. Here `url_normal` is created by `reactive()` and is called "reactive expression". `url_normal()` is a string value containing the concatenated URL. In other words, the `url_normal()` value is the result of the `url_normal` expression.
:::

**Finally**, in your **survey body**, output this `redirect_normal` URL using the `sd_output()` function:

```{r}
# Put this in the desired place of your survey body
sd_output("redirect_normal")
```

After all these are done, your `sd_output()` function will generate a button that redirects you to this link:

```{r}
#| echo: true
#| eval: false
https://www.google.com/?id_a=a123id_b=b234id_c=c345&status=0
```

::: {.callout-tip}

Count-down delay (only applies for **static** redirect):

```{r}
# count-down of 10 seconds
sd_redirect(
  # previous parameters...
  delay  = 10
)
```

Open at a new tab (applies for both **static** and **reactive**):

```{r}
sd_redirect(
  # previous parameters...
  newtab  = TRUE
)
```

If `newtab` is used together with `delay`, the new tab opened after count-down might be blocked by the browser. We've tested it with Safari and Edge. It worked fine on Safari, but Edge blocks it with a pop-up notice.
:::

