---
title: "Survey Components Overview"
---

```{r}
#| eval: true
#| echo: false

library(surveydown)
```

## Overview

Every surveydown survey is composed of a _survey_ and an _app_, defined in two separate files:

- `survey.qmd`: This file is where you define the main survey content of your survey, such as pages, questions, and navigation buttons. It is a standard Quarto document, so you can use markdown (or the RStudio visual editor) to insert text, images, etc. just like you would in any Quarto document. See the [Survey File](survey-file.qmd) page for more details.
- `app.R`: This file is where you define the global settings (libraries, database configuration, etc.) and server logic (e.g., conditional page skipping / question display, etc.) for your survey. It defines a Shiny app, so if you are familiar with Shiny you should feel right at home. See the [App File](app-file.qmd) page for more details.

**These files must be named `survey.qmd` and `app.R`**.
 
They typically look something like this:

::: {.panel-tabset}

## `survey.qmd` file

````{r}
---
format: html
echo: false
warning: false
---

```{r}
library(surveydown)
```

::: {#welcome .sd-page}

# Welcome to our survey!

```{r}
sd_question(
  type  = 'mc',
  id    = 'penguins',
  label = "Which type of penguin do you like the best?",
  option = c(
    'Adélie'    = 'adelie',
    'Chinstrap' = 'chinstrap',
    'Gentoo'    = 'gentoo'
  )
)

sd_next(next_page = 'end')
```

:::

::: {#end .sd-page}

This is the last page of the survey.

:::

````

## `app.R` file

```{r}
library(surveydown)

# sd_db_config()
db <- sd_db_connect()

server <- function(input, output, session) {

  # Conditional skip logic here (skip to page if a condition is TRUE)
  sd_skip_forward()

  # Conditional display logic here (show a question if a condition is TRUE)
  sd_show_if()

  # Main server to control the app
  sd_server(db)
}

shiny::shinyApp(ui = sd_ui(), server = server)
```

:::
