---
title: "Custom Questions"
toc-expand: true
---

```{r}
#| eval: true
#| include: false

library(leaflet)
library(plotly)

# Load state data from tigris - we do this outside of the server
# because we only need to do it once across all sessions
states <- tigris::states(cb = TRUE, resolution = "20m") |>
  dplyr::filter(STUSPS %in% state.abb) |>
  sf::st_transform(4326)
```

If you have an html widget that isn't yet directly supported by surveydown, you can still use it by defining it in the server and using the `sd_question_custom()` function to store the value in the resulting survey data.

## Basic Syntax

Use the `sd_question_custom()` function to create custom questions. The function requires the following arguments:

- `id`: A unique identifier for the question, which will be used as the variable name in the resulting survey data (just like the `id` in `sd_question()`).
- `label`: The label that will be displayed on the question in the survey.
- `output`: The output widget. This needs to be an output function designed for shiny, e.g. `leafletOutput()`, `plotOutput()`, etc.
- `value`: The value to be returned by the question. This must be a reactive value that updates based on user interaction with the `output` widget (e.g. selecting a state on a leaflet map).
- `height` (optional): The height of the widget in pixels, defaults to 400.

Below is the basic syntax for creating a custom question:

```r
sd_question_custom(
  id     = "some_id",
  label  = "Some Label",
  output = "some_output_widget",
  value  = "some_reactive_value"
)
```

To make `sd_question_custom()` work, you need to define the UI and functionality you want in the `server()` function in the `app.R` file. We have two examples that show how to set this up: an interactive map using leaflet, and an interactive plot using plotly.

## `leaflet` map example

See the [demo survey](https://github.com/surveydown-dev/demos/tree/main/leaflet-map) of a leaflet map with the `sd_question_custom` function.

The demo has a leaflet map that looks like the map below. When a user clicks on the map, it turns orange and stores the selected state in the resulting survey response data:

> Which state do you live in?

```{r}
#| eval: true
#| echo: false
#| message: false

# Helper function for modifying the leaflet map layout
map_layout <- function(map, states, selected_state = NULL) {

    # Set the state fill colors
    color <- "lightblue"
    if (!is.null(selected_state)) {
      color <- ifelse(states$NAME == selected_state, "orange", "lightblue")
    }

    # Update the polygons
    addPolygons(
      map,
      data = states,
      fillColor = color,
      weight = 2,
      opacity = 1,
      color = "white",
      fillOpacity = 0.7,
      layerId = states$NAME,
      highlight = highlightOptions(
        weight = 3,
        color = "#666",
        fillOpacity = 0.7,
        bringToFront = TRUE
      ),
      label = ~NAME,
      labelOptions = labelOptions(
        style = list(
          padding = "3px 8px",
          "background-color" = "rgba(255,255,255,0.8)"
        ),
        textsize = "15px"
      )
    )
}

# Create the main leaflet map widget
leaflet(options = leafletOptions(preferCanvas = TRUE)) |>
  addTiles() |>
  setView(lng = -98.5795, lat = 39.8283, zoom = 4) |>
  map_layout(states)
```

## Example of `plotly`

See the [demo survey](https://github.com/surveydown-dev/demos/tree/main/plotly) of a plotly question utilizing the `sd_question_custom` function.

The demo has a plotly scatterplot map that looks like the chart below. When a user clicks on one of the points, it displays the point value and stores the point in the resulting survey response data:

> Click on a point to select it:

```{r}
#| eval: true
#| echo: false
#| message: false

plot_ly(
    mtcars, x = ~wt, y = ~mpg,
    type = "scatter",
    mode = "markers",
    source = "scatter_plot"
) %>%  # Add source identifier
    layout(dragmode = "select")   # Enable point selection
```

