---
title: "Deploying Your Survey"
---

## Preparing to deploy

Before deploying your survey, make sure everything is working locally. You should have done the following:

1. You have locally previewed your `app.R` file by clicking the "Run App" button in RStudio or in your R console running the code `shiny::runApp('app.R')`.
2. You have set up a database connection to store survey responses (e.g. using {{< var supabase >}}) and have already added the database parameters in the `sd_database()` function in your `app.R` file (see the [Store Data](store-data.qmd) page).
3. You have set the password for your survey using `surveydown::sd_set_password("my_password")` using the password you set in the Supabase project (see the [Password](password.qmd) page).

With these steps completed, you are ready to deploy your survey online.

## Deploying to shinyapps.io

Deploying your survey is pretty much the same as deploying any other Shiny app. We recommend using {{< var shinyapps >}} to host your survey as it is designed to work with Shiny Apps, but you can also use [other hosting services](#deploying-to-other-hosting-services).

::: {.callout-note}

You may have deployed other Quarto documents on [Quarto Pub](https://quartopub.com) before, but this site is only for static websites, so you **SHOULD NOT** use Quarto Pub for your survey deployment.

:::

To start using {{< var shinyapps >}}, you'll need to create an account and follow the basic instructions to set up your sub-domain and authorize your IDE. See more information [here](https://docs.posit.co/shinyapps.io/guide/getting_started/).

Once you have your account and sub-domain ready, make sure you have the [rsconnect](https://rstudio.github.io/rsconnect/) package installed. You should have done this while authorizing your IDE, but in case you didn't, you can install it with:

```{r}
install.packages('rsconnect')
```

Then to deploy your survey, run:

```{r}
# Define your app name with the appName parameter
rsconnect::deployApp(appName = "My Survey")
```

That's it! Now you should have your survey site deployed on **shinyapps.io**. Congratulations! 🎉

## Deploying to Other Hosting Services

{{< include chunks/deploy-alts.qmd >}}
