---
title: "Deploying Your Survey"
---

## Preparing to deploy

Before deploying your survey, make sure you have done the following:

1. You have rendered your `survey.qmd` document. 
2. You have locally run your `app.R` file by clicking the "Run App" button in RStudio or in your R console running the `runApp()` command.
3. You have set up a {{< var supabase >}} project and have already added the database parameters in the `sd_database()` function in the setup code chunk (see the [Store Data](store-data.qmd) page for details).
4. You have set the password for your survey using `surveydown::sd_set_password("my_password")`, using the password you set in the Supabase project.
5. You have created an account on {{< var shinyapps >}}, which will be used for survey online deployment.

With these steps completed, you are ready to deploy your survey online.

## Deploying to shinyapps.io

Deploying your survey is pretty much the same as deploying any other Shiny app. 

We recommend using {{< var shinyapps >}} to host your survey as it is designed to work with Shiny Apps, but you can also use other hosting services such as [RStudio Connect](https://docs.posit.co/connect/) or {{< var huggingface >}}. 

::: {.callout-note}

You may have deployed other Quarto documents on [Quarto Pub](https://quartopub.com) before. Note that Quarto Pub is only for static websites, while your survey is an interactive website. Therefore, you **SHOULD NOT** use Quarto Pub for your survey deployment.

:::

To start using shinyapps.io, you'll need to create an account and follow the basic instructions to set up your sub-domain and authorize your IDE. See more information [here](https://docs.posit.co/shinyapps.io/guide/getting_started/).

Once you have your account and your sub-domain ready, it's time to deploy your survey website.

Make sure you have the [rsconnect](https://rstudio.github.io/rsconnect/) {{< fa brands r-project >}} package installed (you should have done this while authorizing your IDE):

```{r}
install.packages('rsconnect')
```

Then, the simplest way to deploy your survey is to run this:

```{r}
rsconnect::deployApp(appName = "Survey Name")
```

The tab name of the HTML page is defined by the `appName` argument. 

That's it! Now you should have your survey site deployed on **shinyapps.io**, with its data stored on **Supabase**. Congratulations!
