---
title: "Getting Started"
---

```{r}
#| echo: false
#| eval: true

library(surveydown)
```

## Overview

In surveydown, surveys are designed using markdown and R code in a single [Quarto]({{< var url_quarto >}}) document (a `.qmd` file). There is no GUI or drag-and-drop interface - simply write plain text (markdown & R code) and boom - you have a survey! 🎉

The [{surveydown}]({{< var url_package >}}) R package provides a set of functions for defining the survey content and configuration options. Each function starts with `sd_` to make them easy to identify.

The platform is based on some basic principles:

- Add content to your survey using markdown text (or in RStudio use the visual editor).
- Define survey questions in R code chunks with the `sd_question()` function.
- Define pages using fences (`:::`), with navigation buttons handled using the `sd_next()` function.
- Add rich functionality to your survey using [config options](config-options.qmd) in the server code chunk.
- Store your respondent data on Supabase by setting the [database](store-data.qmd) in the server code chunk.

This approach ensures a flexible survey platform that is fully reproducible and easy to customize.

The remaining steps on this page will guide you through the process of creating a surveydown survey.

## 1. Install

See the [Installation](installation.qmd) page.

## 2. Start with a template

{{< include chunks/template.qmd >}}

## 3. Add content

See the [Survey Components](survey-components.qmd) page for details on the main components in a surveydown survey. For a quick overview, here's how you add pages and questions:

- Add pages with fences, like this:

```{r}
::: {#page1 .sd-page}

Page 1 content here

:::
```

- Add questions with the `sd_question()` function in code chunks (see the [Question Types](question-types.qmd) page for more on the types of questions supported). For example, here's a multiple choice question:

```{r}
#| echo: fenced

sd_question(
  type  = 'mc',
  id    = 'penguins',
  label = "Which is your favorite type of penguin?",
  option = c(
    'Adélie'    = 'adelie',
    'Chinstrap' = 'chinstrap',
    'Gentoo'    = 'gentoo'
  )
)
```

Output:

```{r}
#| echo: false
#| eval: true

sd_question(
  type  = 'mc',
  id    = 'penguins',
  label = "Which is your favorite type of penguin?",
  option = c(
    'Adélie'    = 'adelie',
    'Chinstrap' = 'chinstrap',
    'Gentoo'    = 'gentoo'
  )
)
```

## 4. Add configuration options

In the server chunk (bottom of the `.qmd` file), add control logic and other configuration options to your survey with the `sd_config()` function. See the [Configuration Options](config-options.qmd) page for more details.

## 5. Setup your database

Also in the server chunk, setup your database with the `sd_database()` function. You can also leave it blank to preview / edit your survey without database connected, or set `pause = TRUE` to disconnect the survey. See the [Store Data](store-data.qmd) page for more details.

## 6. Locally preview

Preview your survey by clicking the "Run Document" button in RStudio or in your terminal running the command:

```{r}
quarto serve survey_file_name.qmd
```

Replacing `survey_file_name.qmd` with the name of your survey `.qmd` file.

## 7. Deploy

Deploy your survey by hosting it on your favorite server, like [shinyapps.io]({{< var url_shinyapps >}}), [huggingface]({{< var url_huggingface >}}), etc. See the [Deployment](deployment.qmd) page for more details.
