---
title: "Getting Started"
---

```{r}
#| echo: false
#| eval: true

library(surveydown)
```

## Overview

{{< include chunks/files.qmd >}}

The [{surveydown}]({{< var url_package >}}) R package provides a set of functions for defining the survey content and configuration options. Each function starts with `sd_` to make them easy to identify.

The platform is based on some basic principles:

- Add content to your `survey.qmd` file using markdown text (or in RStudio use the visual editor).
- Define survey questions in R code chunks with the `sd_question()` function.
- Define pages using fences (`:::`), with navigation buttons handled using the `sd_next()` function.
- Add rich functionality to your survey using [config options](config-options.qmd) in the server function in the `app.R` file.
- Store your respondent data in a database (see [Store Data](store-data.qmd)).

This approach ensures a flexible survey platform that is fully reproducible and easy to customize.

The remaining steps on this page will guide you through the process of creating a surveydown survey.

## 1. Install

See the [Installation](installation.qmd) page.

## 2. Start with a template

See the [Template](template.qmd) page.

## 3. Add survey content in your `survey.qmd` file

See the [Survey Components](survey-components.qmd) page for details on the main components in a surveydown survey. For a quick overview, here's how you add pages and questions:

- Add pages with fences, like this:

```{r}
::: {#page1 .sd-page}

Page 1 content here

:::
```

- Add questions with the `sd_question()` function in code chunks (see the [Question Types](question-types.qmd) page for more on the types of questions supported). For example, here's a multiple choice question:

::: {.panel-tabset}

## Code chunk

```{r}
#| echo: fenced

sd_question(
  type  = 'mc',
  id    = 'penguins',
  label = "Which is your favorite type of penguin?",
  option = c(
    'Adélie'    = 'adelie',
    'Chinstrap' = 'chinstrap',
    'Gentoo'    = 'gentoo'
  )
)
```

## Output

```{r}
#| eval: true
#| echo: false

sd_question(
  type  = 'mc',
  id    = 'penguins',
  label = "Which is your favorite type of penguin?",
  option = c(
    'Adélie'    = 'adelie',
    'Chinstrap' = 'chinstrap',
    'Gentoo'    = 'gentoo'
  )
)
```

:::

## 4. Add configuration options

In the server function in the `app.R` file, add control logic and other configuration options to your survey. See the [Configuration Options](config-options.qmd) page for more details.

## 5. Setup your database

In the global settings at the top of the `app.R` file, setup your database with the `sd_database()` function. You can also leave it blank to preview / edit your survey without database connected, or set `ignore = TRUE` to run the survey without storing data. See the [Store Data](store-data.qmd) page for more details.

## 6. Locally preview

Preview your survey by clicking the "Run App" button in RStudio or in your R console running the `runApp()` command.

## 7. Deploy

Deploy your survey by hosting it on your favorite server, like [shinyapps.io]({{< var url_shinyapps >}}), [huggingface]({{< var url_huggingface >}}), etc. See the [Deployment](deployment.qmd) page for more details.
