---
title: "Getting Started with surveydown"
---

```{r child = here::here("chunks", "intro.qmd")}
```

This introduction covers the basic steps to get started with surveydown.

# 1. Install

See `vignette("install")`

# 2. Start with a template

In the R console, run the following to to setup a template survey:

```{r}
surveydown::sd_create_survey("path/to/folder")
```

This will create a folder located at `"path/to/folder"` with the following files:

- `example.qmd`: a template survey you should edit.
- `example.Rproj`: An RStudio project file (helpful if you're working in RStudio)
- `_extensions`: A folder with the [surveydown Quarto extension](https://github.com/jhelvy/surveydown-ext) needed to make everything work (don't modify this).

# 3. Add content

See `vignette("content")` for details on adding content to your survey, like text, images, etc. with markdown and / or code chunks. As a quick overview:

- Add pages using fences, like this:

```
::: {#page1 .sd-page}

Page 1 content here

:::
```

- Add questions with the `sd_question()` function in code chunks (see `vignette("questions")` for supported question types). For example:

````
`r ''````{r}
sd_question(
  type  = 'mc',
  id    = 'penguins',
  label = "Which is your favorite type of penguin?",
  option = c(
    'Adélie'    = 'adelie',
    'Chinstrap' = 'chinstrap',
    'Gentoo'    = 'gentoo'
  )
)
```
````

# 4. Add configuration options

In the server chunk (bottom of qmd file), add control logic to your survey with the `sd_config()` function. See `vignette("config")`.

# 5. Setup your database

Also in the server chunk, setup your database with the `sd_database()` function, or leave it blank to preview / edit your survey without database connected. See `vignette("database")`.

# 6. Locally preview

Preview your survey by clicking the "Run Document" button in RStudio or in your terminal running the command `quarto serve survey_file_name.qmd`.

# 7. Deploy

Deploy your survey by hosting it on your favorite server, like shinyapps.io, huggingface, etc. See `vignette("deploy")`.
