---
title: "Installation"
---

# Install R & Quarto

You need both: 

- [Install {{< fa brands r-project >}}]({{< var url_rproject >}})
- [Install Quarto]({{< var url_quarto >}})

Also, we recommend working with an IDE. [RStudio](https://www.rstudio.com/categories/rstudio-ide/) is great, and we also like [VS Code](https://code.visualstudio.com) and [Positron](https://github.com/posit-dev/positron).

# Install the {surveydown} R package

The {surveydown} R package is not yet on CRAN, but you can install the development version from GitHub with the following command in your R console:

```{r}
# install.packages("remotes")
remotes::install_github("surveydown-dev/surveydown")
```

# Install the surveydown Quarto Extension

Once you start working on an actual survey, you'll need the [surveydown Quarto extension]({{< var url_extension >}}). 

> **Note**: The extension must be locally installed in a `_extensions` folder in your root survey project folder.

We recommend starting with a template survey, which will automatically add the latest extension for you:

```{r}
surveydown::sd_create_survey("path/to/folder")
```

Otherwise, you can install the extension using the terminal or with the {surveydown} package:

- Using the Terminal:

```{r}
quarto install extension jhelvy/surveydown-ext
```

- Using the {surveydown} R package:

```{r}
surveydown::sd_update_extension()
```

# Version Control

We made some R functions to make it easier to keep your R package and Quarto extension up to date. 

To _check_ if you have the latest R package and Quarto extension installed, run:

```{r}
surveydown::sd_check_versions()
```

To _update_ the R package and Quarto extension to the latest version, run:

```{r}
surveydown::sd_update_surveydown()
```
