---
title: "Installation"
---

# Install R & Quarto

You need both: 

- [Install R](https://CRAN.R-project.org/)
- [Install Quarto](https://quarto.org/docs/get-started/)

Also, we recommend working with an IDE. [RStudio](https://www.rstudio.com/categories/rstudio-ide/) is great, and we also like [VS Code](https://code.visualstudio.com) and [Positron](https://github.com/posit-dev/positron).

# Install the {surveydown} R package

The {surveydown} R package is not yet on CRAN, but you can install the development version from GitHub with the following command in your R console:

```{r}
# install.packages("remotes")
remotes::install_github("jhelvy/surveydown")
```

# Install the surveydown Quarto Extension

Once you start working on an actual survey, you'll need the [surveydown Quarto extension]({{< var url_extension >}}). 

> **Note**: The extension must be locally installed in a `_extensions` folder in your root survey project folder.

We recommend starting with a template survey, which will automatically add the latest extension for you:

```{r}
surveydown::sd_create_survey("path/to/folder")
```

Otherwise, you can install the extension using the *terminal* or with the {surveydown} package:

- Using the *Terminal*:

```{r}
quarto install extension jhelvy/surveydown-ext
```

- Using the {surveydown} R package:

```{r}
surveydown::sd_update_extension()
```

# Version Control

To check if you have the latest R package and Quarto extension versions installed, run:

```{r}
surveydown::sd_check_versions()
```

To update the R package and Quarto extension to the latest version:

```{r}
surveydown::sd_update_surveydown()
```
