---
title: "Setting Your Password"
---

Every surveydown survey requires a password, which is used to communicate with the Supabase database and access the admin page (should you choose to enable it). 

::: {.callout-important}

You should first define your password on Supabase (see [Store Data](database.html#create-a-supabase-project)). Then open your R console to your project folder and set this password with `sd_set_password()`.

:::

## Setting your project password

To set your surveydown password in a given project, run:

```{r}
sd_set_password("your_password")
```

::: {.callout-important}

Do **NOT** put the `sd_set_password()` function in your `app.R` file!

:::

This is designed to be run once in your R console and not again (unless you want to change the password).

## What does this do?

By running `sd_set_password()`, the following happens:

1. A `.Renviron` file is created in your project directory that contains your password stored as an environment variable, e.g., `SUPABASE_PASSWORD=my_password`.
2. If there is a previously defined `SUPABASE_PASSWORD` in `.Renviron`, it will be overwritten, so you are safe to run this function multiple times.
3. A `.gitignore` file is created in your project directory with `.Renviron` in it so it won't be accidentally pushed to GitHub. If a `.gitignore` file already exists, it will be updated to include `.Renviron`.

**The `.Renviron` file must remain in the root project directory - do not manually modify it**

## What is the password used for?

Your password is used for two purposes:

1. Obtaining access to the Supabase database storing your survey data (see [Store Data](store-data.qmd)).
2. Obtaining access to the [admin page](config-options.qmd#admin-page) of your survey.

::: {.callout-important}

Be sure to use the same password when setting up your Supabase database.

:::

## Forgot your password?

To reveal your password, run:

```{r}
sd_show_password()
```

::: {.callout-important}

This is a one-time operation, so you can also run in your R console. You will be prompted to confirm you really want to reveal your password, since it is a sensitive operation.

:::
