---
title: "Setting Your Password"
---

::: {.callout-note}

To avoid confusion, you need to deal with password both on Supabase and surveydown. Firstly, Supabase is where your password is **defined** (see [Store Data](database.html#create-a-supabase-project) for details). Then, on surveydown, you deliver this password using `sd_set_password()`.

:::

::: {.callout-tip}

One Supabase project can have multiple surveydown surveys, but we recommend that you have only one survey for one project, and for each Supabase project, you define a unique password.

:::
## Setting your project password

To set your surveydown password in a given project, run:

```{r}
sd_set_password("your_password")
```

::: {.callout-important}

Do **NOT** put the `sd_set_password()` function in your `.qmd` file!

:::

This is designed to be run once in your R console and not again (unless you want to change the password).

## What does this do?

By running `sd_set_password()`, a `.Renviron` file is created in your project directory that contains your password. This file is used to access your database and write data to it, as well as provide access to the admin page. **This file must remain in the root project directory - do not modify it**.

The call to `sd_set_password()` also adds a `.gitignore` file in your project with `.Renviron` in it, so it won't be accidentally pushed to GitHub. If a `.gitignore` file already exists, it will be updated to include `.Renviron`.

## What is the password used for?

Your password is used for two purposes:

1. Obtaining access to the [admin page](config-options.qmd#admin-page) of your survey.
2. Obtaining access to the Supabase database storing your survey data (see [Store Data](store-data.qmd)).

::: {.callout-important}

Be sure to use the same password when setting up your supabase database.

:::

## Forgot your password?

To reveal your password, run:

```{r}
sd_show_password()
```

::: {.callout-important}

This is a one-time operation, so you can also run in your R console instead of the `.qmd` file. You will be prompted to confirm you really want to reveal your password, since it is a sensitive operation.

In short: **SECURE YOUR PASSWORD!**

:::