---
title: "Question Types"
toc-expand: true
---

```{r}
#| eval: true
#| echo: false

library(surveydown)
```

# Introduction to `sd_question()`

All questions in a surveydown survey are created using the `sd_question()` function.

The function requires the following three arguments:

- `id`: A unique identifier for the question, which will be used as the variable name in the resulting survey data.
- `label`: The label that will be displayed on the question in the survey.
- `type`: The type of question.

Many question types also require an `option` argument, which is a named vector of options for the question (e.g. for multiple choice questions).

The function also includes many other optional arguments that can be used to customize the question appearance and behavior. See the [package documentation](https://pkg.surveydown.org/reference/sd_question) for more details.

# Question Types

The following question types are supported by `{surveydown}`:

- `text`: Single line open text input.
- `textarea`: Multiple line open text input.
- `numeric`: Single line numeric text input.
- `mc`: Multiple choice with a single select option (radio buttons).
- `mc_buttons`: Same as `mc` but as a "button" style instead of radio buttons.
- `mc_multiple`: Multiple choice with mutliple select options (check boxes).
- `mc_multiple_buttons`: Same as `mc_multiple` but as a "button" style instead of check boxes.
- `select`: Select a choice from a drop down menu.
- `slider`: (Under construction) Slider to select discrete categories or continuous numbers.
- `date`: Select a date from a calendar widget.
- `daterange`: Select two dates from calendar widgets (e.g. begin and end dates).

Below are examples of each question type.

## `text`

Use `type = 'text'` to specify a text input type question.

Code chunk:

```{r}
#| echo: fenced

sd_question(
  type  = "text",
  id    = "word",
  label = "Write a word:"
)
```

Output:

```{r}
#| eval: true
#| echo: false

sd_question(
  type  = "text",
  id    = "word",
  label = "Write a word:"
)
```

## `textarea`

Use `type = 'textarea'` to specify a text area input type question.

Code chunk:

```{r}
#| echo: fenced

sd_question(
  type  = "textarea",
  id    = "paragraph",
  label = "Write a paragraph:"
)
```

Output:

```{r}
#| eval: true
#| echo: false

sd_question(
  type  = "textarea",
  id    = "paragraph",
  label = "Write a paragraph:"
)
```

## `numeric`

Use `type = 'numeric'` to specify a numeric input type.

Code chunk:

```{r}
#| echo: fenced

sd_question(
  type  = 'numeric',
  id    = 'age',
  label = "What's your age?"
)
```

Output:

```{r}
#| eval: true
#| echo: false

sd_question(
  type  = 'numeric',
  id    = 'age',
  label = "What's your age?"
)
```

## `mc`

Use `type = 'mc'` to specify a multiple choice type question with a single choice option.

Code chunk:

```{r}
#| echo: fenced

sd_question(
  type   = 'mc',
  id     = 'artist',
  label  = "Which artist do you prefer most from this list?",
  option = c(
    "Taylor Swift" = "taylor_swift",
    "Beyoncé"      = "beyonce",
    "Adele"        = "adele",
    "Rihanna"      = "rihanna",
    "Lady Gaga"    = "ladygaga",
    "Ed Sheeran"   = "ed_sheeran",
    "Drake"        = "drake"
  )
)
```

Output:

```{r}
#| eval: true
#| echo: false

sd_question(
  type   = 'mc',
  id     = 'artist',
  label  = "Which artist do you prefer most from this list?",
  option = c(
    "Taylor Swift" = "taylor_swift",
    "Beyoncé"      = "beyonce",
    "Adele"        = "adele",
    "Rihanna"      = "rihanna",
    "Lady Gaga"    = "ladygaga",
    "Ed Sheeran"   = "ed_sheeran",
    "Drake"        = "drake"
  )
)
```

## `mc_buttons`

Use `type = 'mc_buttons'` to generate the button version of `mc`.

Code chunk:

```{r}
#| echo: fenced

sd_question(
  type   = 'mc_buttons',
  id     = 'fruit',
  label  = "Which fruit do you prefer most from this list?",
  option = c(
    "Apple"      = "apple",
    "Banana"     = "banana",
    "Pear"       = "pear",
    "Strawberry" = "strawberry",
    "Grape"      = "grape",
    "Mango"      = "mango",
    "Watermelon" = "watermelon"
  )
)
```

Output:

```{r}
#| eval: true
#| echo: false

sd_question(
  type   = 'mc_buttons',
  id     = 'fruit',
  label  = "Which fruit do you prefer most from this list?",
  option = c(
    "Apple"      = "apple",
    "Banana"     = "banana",
    "Pear"       = "pear",
    "Strawberry" = "strawberry",
    "Grape"      = "grape",
    "Mango"      = "mango",
    "Watermelon" = "watermelon"
  )
)
```

## `mc_multiple`

Use `type = 'mc_multiple'` to specify a multiple choice type question with multiple selection enabled.

Code chunk:

```{r}
#| echo: fenced

sd_question(
  type  = 'mc_multiple',
  id    = 'swift',
  label = "What are your favorite Taylor Swift albums (select all that apply)?",
  option = c(
    "Taylor Swift (2006)" = "taylor_swift",
    "Fearless (2008)"     = "fearless",
    "Speak Now (2010)"    = "speak_now",
    "Red (2012)"          = "red",
    "1989 (2014)"         = "1989",
    "Reputation (2017)"   = "reputation",
    "Lover (2019)"        = "lover",
    "Folklore (2020)"     = "folklore",
    "Evermore (2020)"     = "evermore",
    "Midnights (2022)"    = "midnights"
  )
)
```

Output:

```{r}
#| eval: true
#| echo: false

sd_question(
  type  = 'mc_multiple',
  id    = 'swift',
  label = "What are your favorite Taylor Swift albums (select all that apply)?",
  option = c(
    "Taylor Swift (2006)" = "taylor_swift",
    "Fearless (2008)"     = "fearless",
    "Speak Now (2010)"    = "speak_now",
    "Red (2012)"          = "red",
    "1989 (2014)"         = "1989",
    "Reputation (2017)"   = "reputation",
    "Lover (2019)"        = "lover",
    "Folklore (2020)"     = "folklore",
    "Evermore (2020)"     = "evermore",
    "Midnights (2022)"    = "midnights"
  )
)
```

## `mc_multiple_buttons`

Use `type = 'mc_multiple_buttons'` to generate the button version of `mc_multiple`.

Code chunk:

```{r}
#| echo: fenced

sd_question(
  type  = 'mc_multiple_buttons',
  id    = 'michael_jackson',
  label = "Which are your favorite Michael Jackson songs (select all that apply)?",
  option = c(
    "Thriller (1982)"          = "thriller",
    "Billie Jean (1982)"       = "billie_jean",
    "Beat It (1982)"           = "beat_it",
    "Man in the Mirror (1987)" = "man_in_the_mirror",
    "Smooth Criminal (1987)"   = "smooth_criminal",
    "Black or White (1991)"    = "black_or_white",
    "Bad (1987)"               = "bad",
    "Human Nature (1982)"      = "human_nature"
  )
)
```

Output:

```{r}
#| eval: true
#| echo: false

sd_question(
  type  = 'mc_multiple_buttons',
  id    = 'michael_jackson',
  label = "Which are your favorite Michael Jackson songs (select all that apply)?",
  option = c(
    "Thriller (1982)"          = "thriller",
    "Billie Jean (1982)"       = "billie_jean",
    "Beat It (1982)"           = "beat_it",
    "Man in the Mirror (1987)" = "man_in_the_mirror",
    "Smooth Criminal (1987)"   = "smooth_criminal",
    "Black or White (1991)"    = "black_or_white",
    "Bad (1987)"               = "bad",
    "Human Nature (1982)"      = "human_nature"
  )
)
```

## `select`

Use `type = 'select'` to specify a drop down select type question.

Code chunk:

```{r}
#| echo: fenced

sd_question(
  type  = 'select',
  id    = 'education',
  label = "What is the highest level of education you have attained?",
  option = c(
    "Did not attend high school" = "hs_no",
    "Some high school"           = "hs_some",
    "High school graduate"       = "hs_grad",
    "Some college"               = "college_some",
    "College"                    = "college_grad",
    "Graduate Work"              = "grad",
    "Prefer not to say"          = "no_response"
  )
)
```

Output:

```{r}
#| eval: true
#| echo: false

sd_question(
  type  = 'select',
  id    = 'education',
  label = "What is the highest level of education you have attained?",
  option = c(
    "Did not attend high school" = "hs_no",
    "Some high school"           = "hs_some",
    "High school graduate"       = "hs_grad",
    "Some college"               = "college_some",
    "College"                    = "college_grad",
    "Graduate Work"              = "grad",
    "Prefer not to say"          = "no_response"
  )
)
```

## `slider` (Under Construction)

Due to compatibility issues, the `slider` type is still under construction.

## `date`

Use `type = 'date'` to specify a date input type.

Code chunk:

```{r}
#| echo: fenced

sd_question(
  type  = 'date',
  id    = 'dob',
  label = "What is your date of birth?"
)
```

Output:

```{r}
#| eval: true
#| echo: false

sd_question(
  type  = 'date',
  id    = 'dob',
  label = "What is your date of birth?"
)
```

The date value will be today's date by default. Upon clicking on the text box, you are provided with a date dialog box to choose date from.

::: {.callout-note}

The dialog box will only appear in a Shiny environment, hence why it is not appearing when clicked here.

:::

## `daterange`

Use `type = 'daterange'` to specify a date range input type.

Code chunk:

```{r}
#| echo: fenced

sd_question(
  type  = 'daterange',
  id    = 'hs_date',
  label = "When did you start and finish high school?"
)
```

Output:

```{r}
#| eval: true
#| echo: false

sd_question(
  type  = 'daterange',
  id    = 'hs_date',
  label = "When did you start and finish high school?"
)
```

::: {.callout-note}

The dialog box will only appear in a Shiny environment, hence why it is not appearing when clicked here.

:::
