---
title: "System Translations"
---

As of v0.4.2, {surveydown} supports the ability to set the system language for all system messages (i.e. text elements such as button labels and warnings) and to provide the language specific format for [date type questions](question-types.qmd#date). You can also customize the system messages by providing a `translations.yml` file in your root project directory.

## Setting the Language

To set the language of your survey, you can use the `language` argument in the `sd_server()` function in your `app.R` file. For example, to set the survey language to Spanish, you would write:

```{r}
server <- function(input, output, session) {
  sd_server(
    language = "es"
  )
}
```

The `language` argument accepts any of the following [language codes](#supported-language-codes):

```{r}
"ar", "az", "bg", "bs", "ca", "cs", "cy", "da", "de", "el", "en",
"en-AU", "en-GB", "eo", "es", "et", "eu", "fa", "fi", "fo", "fr-CH",
"fr", "gl", "he", "hr", "hu", "hy", "id", "is", "it-CH", "it", "ja",
"ka", "kh", "kk", "ko", "kr", "lt", "lv", "me", "mk", "mn", "ms", "nb",
"nl-BE", "nl", "no", "pl", "pt-BR", "pt", "ro", "rs-latin", "rs", "ru",
"sk", "sl", "sq", "sr-latin", "sr", "sv", "sw", "th", "tr", "uk", "vi",
"zh-CN", "zh-TW"
```

In addition, {surveydown} comes with built-in translations for the following languages:

- English (`"en"`, default)
- German (`"de"`)
- Spanish (`"es"`)
- French (`"fr"`)
- Italian (`"it"`)

These translations cover all system messages used in surveydown.

## Custom Messages

If you want to customize the system translations or provide translations in a language that is not yet supported, you can create a `translations.yml` file in the root folder of your survey project using the following function:

```{r}
surveydown::sd_create_translations(language = 'en')
```

This will create a file named `translations.yml` in your root project directory with the language set to whatever you provided as the `language` argument. You can modify any of the system messages in this file to be used in your survey. 

Below is an example of the `translations.yml` file for all default English messages. The keys represent the system message identifiers, and the values are the translated messages:

```yaml
en:
  cancel: Cancel
  confirm_exit: Confirm Exit
  sure_exit: Are you sure you want to exit the survey?
  submit_exit: Submit and Exit
  warning: Warning
  required: Please answer all required questions before proceeding.
  rating_title: Before you go...
  rating_text: 'Rate your survey experience:'
  rating_scale: from 1-poor to 5-excellent
  next: Next
  exit: Exit Survey
  close_tab: Please close this tab manually to exit the survey.
  choose_option: Choose an option...
  click: Click here
  redirect: Redirecting in
  seconds: seconds
  new_tab: Opens in a new tab
  redirect_error: 'Error: This text won''t trigger any redirection...'
```

When you run your survey, {surveydown} will detect the `translations.yml` file in your root folder and use it to override the default translations. If you provide translations for a language not supported by default, you can use the `language` argument in `sd_server()` to set your custom language code.

For example, suppose you added translations for Portuguese in your `translations.yml`:

```yaml
pt:
  next: 'Próximo'
  exit: 'Sair da Pesquisa'
  # ... other messages
```

In your `app.R` file, set the language to Portuguese:

```{r}
server <- function(input, output, session) {
  sd_server(
    language = "pt"
  )
}
```

It is also possible to have multiple translations for different translations in a single `translations.yml` file and simply select the required language one within the `sd_server()` function with the matching country code.

## Overriding Specific Messages

You don't have to provide translations for all messages. If you only want to change specific messages, you can provide translations only for those messages, and {surveydown} will use the default translations for any missing messages.

For example, to customize only all "Next" button label in English, you could include the following in your `translations.yml` file:

```yaml
en:
  next: 'Continue'
```

Since English is the default language, the language doesn't even have to be defined in the `sd_server()` function in this case.

## Button Defaults

The translations also apply to default labels in `sd_next()`, `sd_close()`, and `sd_redirect()`. For example, if you use `sd_next()` without specifying a `label`, it will use the translated label based on the chosen language.

```{r}
# In your survey.qmd file

# This will display "Weiter" if language is set to "de"
sd_next()
```

However, it is still possible to manually adjust the translations in the mentioned functions inside the `survey.qmd` file using the `label` argument for each case.

## Notes

- When providing custom translations, make sure the keys match exactly the expected message identifiers.
- If you create the `translations.yml` file manually, please note that each text element containing a `:` character must be written in quotation marks and the last line must contain a line break.
- If you provide an invalid language code in `language`, {surveydown} will fall back to English.
- If a message is not translated in your `translations.yml` file, {surveydown} will use the default translation for that message.

## Supported Language Codes

Surveydown uses the languages supported by Shiny's `dateInput()` for date inputs. The full list of supported language codes can be found in the documentation for [`shiny::dateInput()`](https://shiny.posit.co/r/reference/shiny/1.7.0/dateinput). When setting the `language` argument, make sure to use a valid language code.
