---
title: "Start with a Template"
---

We recommend starting with a template to build your surveydown survey.

In the R console, run the following to to setup a template survey:

```{r}
surveydown::sd_create_survey("path/to/folder")
```

This will create a folder located at `"path/to/folder"` with the following files:

- `survey.qmd`: a template survey where you define your survey content (text, questions, images, etc.).
- `app.R`: a template app file where you define your database connection, survey configuration options, and any other server logic.
- `survey.Rproj`: An RStudio project file (helpful if you're working in RStudio)

::: {.callout-important}

Every survey created with {surveydown} should be in its own separate project folder.

:::

## Single vs Multiple File Structure

The `app.R` file defines the Shiny app, which by default is a single file. If you wish to define your app using separate files, you can do so by setting the `structure` argument to `"multi"` in the `sd_create_survey()` function. This will create a template with four files:

```{r}
surveydown::sd_create_survey("path/to/folder", structure = "multi")
```

- `survey.qmd`: a template survey where you define your survey content (text, questions, images, etc.).
- `ui.R`: The main UI file for the Shiny app. It will only contain the `sd_ui()` function.
- `server.R`: a template server file where you define the server logic of the app.
- `global.R`: a template global file where you define any global variables, libraries, or functions.
