---
title: "Fetching Data"
---

## Static Fetching

Once you have your database properly set up and your survey is live, you can fetch the data using the `sd_get_data()` function. 

To do so, you have to first create a database connection using the `sd_db_connect()` function:

```{r}
db <- sd_db_connect()
```

Once you have created a successful database connection, you need to use the connection to fetch the data, like this:

```{r}
data <- sd_get_data(db)
```

Here the `data` object will be a data.frame with the data from your database table.

## Live-polling Data

::: {.callout-tip}

Run the [live-polling template](https://github.com/surveydown-dev/template_live_polling) of a reactive plot for an example of live-polling data.

:::

The above example is intended for an analysis context, where all you want to do is to fetch the latest data from your database. However, you may also want to fetch the data in a reactive context (i.e. in your survey). For example, perhaps you want to show respondents a bar chart of the most common responses to a question. The `sd_get_data()` function is compatible with reactive fetching as well - just call it inside the `server()` function in your **app.R** file:

```{r}
server <- function(input, output, session) {

  data <- sd_get_data(db)

  sd_server()

}
```

Now the `data` object will be a reactive expression, not a data.frame. To use it, you have to call it with `()` to get the latest data as a data.frame. For example:

```{r}
server <- function(input, output, session) {

  data <- sd_get_data(db, refresh_interval = 5)

  output$my_plot <- renderPlot({  
    my_data <- data()
    # insert code here to make a plot with my_data
  })

  sd_server()

}
```

Here the `my_data` object will be a data.frame with the latest data from your database, refreshed every 5 sections according to the `refresh_interval` parameter (5 seconds is the default value), which you could then use to make a plot. 

In this example, the `output$my_plot` object will be a plot, which you can then display in your survey by placing `plotOutput("my_plot")` in a code chunk somewhere in your **survey.qmd** file.

See the [Reactivity](reactivity.html) page for more information on how to use reactivity in your survey.
