---
title: "Getting Started"
eval: false
css: ../css/surveydown.css
---

```{r}
#| echo: false
#| eval: true

library(surveydown)
shiny::shinyOptions(bootstrapTheme = bslib::bs_theme(version = 5L))
```

This page covers everything you need to get started making a survey with surveydown. 

## Overview

{{< include ../chunks/files.qmd >}}

The [**surveydown**]({{< var url_package >}}) R package provides a set of functions for defining the survey content and configuration options. Each function starts with `sd_` to make them easy to identify.

The platform is based on some basic principles:

- All content is defined and controlled in code, making it fully reproducible and version controllable.
- Survey content is defined inside a [Quarto](https://quarto.org/) document, allowing for rich content combining Markdown syntax and code chunks.
- Each survey is rendered as a single Shiny application, allowing for interactivity and dynamic behavior.
- Data storage is flexible and managed independently of where the survey app is hosted, providing flexibility and control over all data handling and storage.

The remaining steps on this page will guide you through the process of creating a surveydown survey.

## 1. Install

### Install R & Quarto

You need both: 

- Install {{< var rproject >}}
- Install {{< var quarto >}}

We also recommend working with an IDE that has good support for R, Quarto, and Shiny.

{{< var rstudio_ide >}} is great, and we also like {{< var vscode >}} and {{< var positron >}}.

### Install the surveydown R package

You can install surveydown from CRAN in your R console:

```{r}
install.packages("surveydown")
```

or you can install the development version from [GitHub](https://github.com/surveydown-dev/surveydown):

```{r}
# install.packages("pak")
pak::pak("surveydown-dev/surveydown")
```

Load the package with:

```{r}
library(surveydown)
```

You can also check which version you have installed:

```{r}
surveydown::sd_version()
```

## 2. Start with a template

::: {.callout-note}

Every survey created with surveydown should be in its own separate project folder.

:::

We recommend starting with a template to build your first surveydown survey. In the R console, run the following to to setup a generic template survey:

```{r}
surveydown::sd_create_survey()
```

You can also specify where you want the template survey to be created using the `path` argument, and you can also specify which template to use with the `template` argument, like this:

```{r}
surveydown::sd_create_survey(
  path = "path/to/folder",
  template = "question_types"
)
```

The default is `template = "default"`, but you can specify other templates. See the [Templates](../templates.html) page for an overview of all currently available templates.

## 3. Add survey content

Survey content is edited in the **survey.qmd** file. See the [Basic Components](basic-components.qmd) page for more details on how to add more content to a surveydown survey. At a minimum, you can add pages and questions like this:

- Add pages with the shorthand syntax (details [here](page-navigation.html#defining-pages)), like this:

```{r}
--- page1

Page 1 content here

--- page2

Page 2 content here
```

- Add questions with the `sd_question()` function in code chunks (see the [Question Types](question-types.html) page for more on the types of questions supported). For example, here's a multiple choice question:

::: {.panel-tabset}

## Code chunk

```{r}
#| echo: fenced

sd_question(
  type = 'mc',
  id = 'penguins',
  label = "Which is your favorite type of penguin?",
  option = c(
    'Adélie' = 'adelie',
    'Chinstrap' = 'chinstrap',
    'Gentoo' = 'gentoo'
  )
)
```

## Output

```{r}
#| eval: true
#| echo: false

sd_question(
  type = 'mc',
  id = 'penguins',
  label = "Which is your favorite type of penguin?",
  option = c(
    'Adélie' = 'adelie',
    'Chinstrap' = 'chinstrap',
    'Gentoo' = 'gentoo'
  )
)
```

:::

## 4. Add control options

In the `server()` function in the **app.R** file, add rich functionality to your survey using a variety of [Survey Settings](survey-settings.html) or [Conditional Logic](conditional-logic.html).

## 5. Setup your database connection

Setup your database connection using the `sd_db_config()` function. Once your configuration credentials are created (they get saved in a `.env` file), make a connection to your database using the `sd_db_connect()` function in the global settings at the top of the **app.R** file. You can also use `sd_db_connect(ignore = TRUE)` to run the survey without storing data. See the [Storing Data](storing-data.qmd) page for more details.

## 6. Locally preview

Preview your survey by clicking the "Run App" button in RStudio or in your R console running the `runApp()` command.

## 7. Deploy

Deploy your survey by hosting it on your favorite server, like {{< var shinyapps >}}, {{< var huggingface >}}, [Posit Connect Cloud](https://connect.posit.cloud/), [Heroku](https://www.heroku.com/), etc. See the [Deployment](deployment.html) page for more details.
